---
title: "Homework 2"
author: Yanhao Li
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r r_MT}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data! For 2018 and 2017.

```{r r_P2018_2017}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation. 

```{r c_ap}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the NYC Transit dataset.

```{r r_c_NYC}
NYC_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:vending, -exit_only, ada) %>% 
  mutate(
    entry = ifelse(entry %in% c("YES"), TRUE, FALSE)
    )
```

NYC Transit dataset contains information from subway system in NYC. It includes information about line, station, route, entrance, entry/exit status, vending, staff, and ADA. 

So far, I have cleaned the variables' names by clean_names, retained all varibales I want by select, and converted entry variable from character to logical variable by mutate. 

The resulting dataset has `r nrow(NYC_df)` rows and `r ncol(NYC_df)` columns. 

These data have not been tidy yet. For example, route 1 to 11 occupy 11 columns. If we could arrange them, dataset will more likely to be tidy.

Questions answered using these data;

* There are `r nrow(distinct(NYC_df, station_name, line))` distinct stations. 

* There are `r filter(NYC_df, ada == "TRUE") %>% distinct(station_name, line) %>% nrow()` stations as ADA compliant.

* `r filter(NYC_df, vending == "NO", entry == "TRUE") %>% nrow() / filter(NYC_df, vending == "NO") %>% nrow()` of station entrances/exits without vending allow entrance.

Reformat the NYC Transit dataset

```{r r_f_NYC}
re_NYC = NYC_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_prefix = "route",
    names_to = "route_number",
    values_to = "route_name"
  )
```

`r filter(re_NYC, route_name == "A") %>% distinct(station_name, line) %>% nrow()` distinct stations serve the A train.

Of the stations that serve the A train, `r filter(re_NYC, route_name == "A", ada == "TRUE") %>% distinct(station_name, line) %>% nrow()` stations are ADA compliant.

## Problem 3

Read and clean the pols-month dataset

```{r r_c_pm}
pm_df = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    mon,
    into = c("year", "month", "day")
  ) %>% 
  mutate(
    month = recode(
      month, 
      "01" = "jan",
      "02" = "feb",
      "03" = "mar",
      "04" = "apr",
      "05" = "may",
      "06" = "jun",
      "07" = "jul",
      "08" = "aug",
      "09" = "sep",
      "10" = "oct",
      "11" = "nov",
      "12" = "dec")
    ) %>% 
  mutate(
    president = recode(prez_gop, "0" = "democratic", "1" = "republican", "2" = "republican")
  ) %>% 
  select(-prez_gop, -prez_dem, -day)
```

Read and clean the snp dataset

```{r r_c_snp}
snp_df = 
  read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(
    date,
    into = c("month", "day", "year")
  ) %>% 
  mutate(
    month = recode(
      month, 
      "1" = "jan",
      "2" = "feb",
      "3" = "mar",
      "4" = "apr",
      "5" = "may",
      "6" = "jun",
      "7" = "jul",
      "8" = "aug",
      "9" = "sep",
      "10" = "oct",
      "11" = "nov",
      "12" = "dec")
    ) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```

Read and clean unemployment dataset

```{r r_c_unem}
unem_df = 
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "umemployment"
  ) %>% 
  mutate(
    year = as.character(year)
  )
```

Merge three datasets

```{r merge}
pols_snp_unem = 
  left_join(pm_df, snp_df, by = c("year", "month")) %>% 
  left_join(unem_df, by = c("year", "month"))
```

Pols_month dataset has `r nrow(pm_df)` rows and `r ncol(pm_df)` columns. The range of years is from `r max(pull(pm_df, year))` to `r min(pull(pm_df, year))`. It contains information about number of national politicians who are democratic or republican at any given time. It includes key variables named as `r colnames(pm_df)`.

Snp dataset has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. The range of years is from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. It contains information about Standard & Poor’s stock market index, which is often used as a representative measure of stock market as a whole. It includes key variables named as `r colnames(snp_df)`.

Unemployment dataset has `r nrow(unem_df)` rows and `r ncol(unem_df)` columns. The range of years is from `r min(pull(unem_df, year))` to `r max(pull(unem_df, year))`. It contains information about unemployment rate at any given month. It includes key variables named as `r colnames(unem_df)`.

Resulting dataset has `r nrow(pols_snp_unem)` rows and `r ncol(pols_snp_unem)` columns. The range of years is from `r min(pull(pols_snp_unem, year))` to `r max(pull(pols_snp_unem, year))`. It contains information merged from three datasets above. It includes key variables named as `r colnames(pols_snp_unem)`.